We present Kronos, a modular chess engine and benchmarking framework implemented in Node.js and JavaScript, designed to empirically evaluate the execution behavior of classical search heuristics within managed, garbage-collected runtime environments. Standard chess engine implementations favor native compile targets like C/C++ to avoid execution overhead and garbage collection pauses. We observe that through targeted optimization of board representations, strict object-reuse patterns, and modular search configurations, stable search throughput can be achieved on dynamic runtimes. Our architecture decouples the React-based user interface control plane from background search Web Workers executing the search logic. The engine implements standard search enhancements including Principal Variation Search (PVS), Null Move Pruning (NMP), and Late Move Reductions (LMR), backed by a size-bounded, Zobrist-keyed Transposition Table. Using a parallel tournament framework with color-flipped paired matches, our measurements indicate that these combined heuristics compress the search space compared to baseline alpha-beta search without ordering enhancements. Under Sequential Probability Ratio Testing (SPRT), quiescence search was the only enhancement accepted as demonstrating a statistically significant playing-strength improvement, while the remaining heuristics produced node-count reductions but inconclusive playing-strength evidence at the tested sample sizes. Evaluating relative playing strength across configurations yields a peak internal self-play rating of 1,400 Elo at search depth 6, after which worker latency timeouts degrade performance. External calibration matches against depth-limited Stockfish 18 reference benchmarks report a matchup differential of $-215 \pm 29$ Elo at depth 3; we note that depth-limited Stockfish degrades non-linearly due to its NNUE evaluation architecture (see Section~\ref{sec:rating_conventions}), meaning these matchup differentials reflect relative performance bounds under dynamic runtime constraints rather than absolute anchors. These outcomes suggest that carefully structured dynamic runtimes can support stable, real-time game-tree search without performance degradation from memory sweeps.
