\chapter{Engine Configuration Parameters}
\label{app:configs}

This appendix contains sample configuration schemas, opening FEN libraries, Stockfish calibration models, and raw runtime logs.


\section{Search Engine Configuration Schema}
The engine variant behaviors are defined using a JSON configuration file. Below is the configuration spec utilized by the tournament runner to initialize each engine instance:
% Validated and corrected JSON schema configuration
\begin{verbatim}
{
  "search": {
    "useAlphaBeta": true,
    "useIterativeDeepening": true,
    "useMoveOrdering": true,
    "useTranspositionTable": true,
    "useQuiescence": true,
    "usePVS": true,
    "useLMR": true,
    "useNMP": true
  },
  "evaluation": {
    "usePieceSquareTables": true,
    "weights": {
      "pawn": 100,
      "knight": 320,
      "bishop": 330,
      "rook": 500,
      "queen": 900,
      "king": 20000
    },
    "taperedEndgameThreshold": 1600
  },
  "cache": {
    "maxTranspositionTableSize": 500000
  }
}
\end{verbatim}

\section{Opening Book Sample FEN Seeds}
To evaluate playing strength without bias, games are seeded with positions from a randomized opening library. Below is a selection of the baseline opening FENs used:
\begin{enumerate}
    \item \texttt{rnbqkbnr/pppppppp/8/8/4P3/8/PPPP1PPP/RNBQKBNR b KQkq e3 0 1} (King's Pawn Opening)
    \item \texttt{rnbqkbnr/pppppppp/8/8/3P4/8/PPP1PPP/RNBQKBNR b KQkq d3 0 1} (Queen's Pawn Opening)
    \item \texttt{rnbqkbnr/pppp1ppp/8/4p3/4P3/8/PPPP1PPP/RNBQKBNR w KQkq e6 0 2} (Open Game)
    \item \texttt{r1bqkbnr/pppp1ppp/2n5/4p3/4P3/5N2/PPPP1PPP/RNBQKB1R w KQkq - 2 3} (King's Knight Opening)
    \item \texttt{r1bqkbnr/pppp1ppp/2n5/1B2p3/4P3/5N2/PPPP1PPP/RNBQK2R b KQkq - 3 3} (Ruy Lopez)
\end{enumerate}

\chapter{Stockfish Calibration Notes}
\label{app:sf_notes}

Extrapolated playing strength projections for deeper Stockfish configurations (Depths 4--8) are intentionally omitted from this work. Engine strength scaling is strictly non-linear and subject to diminishing returns, making linear extrapolation from shallow-depth calibration data scientifically invalid. Furthermore, as noted in Section~\ref{ch:experimental_results}, depth-limited Stockfish's NNUE evaluation architecture degrades non-linearly, rendering even the measured shallow-depth matchup differentials unsuitable as anchors for extrapolation. Future work requiring deeper calibration should conduct direct head-to-head matches at each target depth.


\chapter{Raw Telemetry Logs}
\label{app:telemetry}
We provide a sample telemetry trace recorded during calibration matches to document V8 runtime memory behavior:

\begin{verbatim}
[Kronos Profiler] Game 5 Complete.
[Memory Profile] RSS: 75.26 MB | Heap Total: 45.26 MB | Heap Used: 26.60 MB
[Cache Telemetry] Transposition Table Size: 48 entries | Purging.
[Kronos Profiler] Game 10 Complete.
[Memory Profile] RSS: 77.31 MB | Heap Total: 46.76 MB | Heap Used: 24.98 MB
[Cache Telemetry] Transposition Table Size: 43 entries | Purging.
\end{verbatim}

\chapter{Reproducibility Checklist}
\label{app:checklist}

To ensure complete empirical transparency and support future verification studies, we provide a systems reproducibility checklist for the Kronos benchmarking suite.

\begin{table}[htbp]
\centering
\caption[Reproducibility Configuration Checklist]{Reproducibility Configuration Checklist. Item indicates the hardware or software parameter; Value / Source lists the specific version, hardware model, or source identifier used to execute the experiments.}
\label{tab:reproducibility_checklist}

\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Item} & \textbf{Value / Source} \\
\midrule
Code Repository & \url{https://github.com/piyush2180/kronos} \\
Commit Hash & \texttt{53389c710f974b8e} \\
Runtime Environment & Node.js v20.11.0 (Google V8 Engine) \\
Host Hardware (CPU) & Intel Core i5-1235U (10 Cores, 12 Threads) \\
Host Hardware (RAM) & 16 GB RAM (16,384 MB) \\
Host OS            & Windows 11 Home Single Language (Build 26100) \\
PRNG Determinism Seed & \texttt{42} (fixed inside LCG shuffle) \\
Opening Book Library & 50 balanced positions (\texttt{openings.json}) \\
SPRT Error Bounds & $\alpha = 0.05$, $\beta = 0.05$ \\
Fitted Bradley-Terry & Anchored to Baseline Minimax (v1.0) = 1000 \\
\bottomrule
\end{tabularx}
\end{table}


