\section*{Executive Summary}
\addcontentsline{toc}{chapter}{Executive Summary}

This project establishes that high-performance, real-time decision engines can run successfully inside web browsers and lightweight server runtimes---environments traditionally avoided by systems engineers due to execution overhead and unpredictable memory cleaning pauses. We designed and built \textbf{Kronos}, an open-source, web-deployable chess engine and benchmarking workstation. Kronos serves as a proof of concept showing that dynamic runtimes (such as Google V8) can execute latency-sensitive search algorithms efficiently, making systems research highly accessible to students and developers without requiring specialized desktop compilers or complex local setups.

\subsection*{Why This Project Matters}
Typically, chess engines are written in native compiled languages like C, C++, or Rust to squeeze every drop of CPU performance and maintain complete control over memory. Interpreted and managed environments like JavaScript are notorious for background ``garbage collection'' sweeps, which periodically freeze the execution thread to clean up unused memory objects, rendering them unsuitable for real-time game search. 

Kronos proves that these performance bottlenecks are not language-inherent. By designing search paths that completely bypass memory allocations on the active search path (using in-place object reuse and size-bounded caching) and decoupling search logic into background processing threads, we achieved smooth, predictable search execution directly in the browser. 

\subsection*{Core Contributions}
This research delivers three major components:
\begin{enumerate}
    \item \textbf{A Web-Based Chess Workstation:} A premium React-based dashboard featuring an interactive play interface, real-time search telemetry, and local puzzle-solving capabilities.
    \item \textbf{A Standardized Research Laboratory:} A CLI-driven benchmarking pipeline that automates self-play tournaments, computes relative playing strength, and verifies statistical significance from any command line.
    \item \textbf{Empirical Architectural Insights:} A detailed performance study validating the exact execution savings of classical search heuristics, establishing how move sorting, deep-tree pruning, and transposition tables interact under dynamic runtime constraints.
\end{enumerate}

Ultimately, Kronos demonstrates that web technologies can host complex, resource-intensive artificial intelligence applications without sacrificing user interface responsiveness or runtime stability, lowering the barrier of entry for systems research.
