\chapter{Experimental Results}
\label{ch:experimental_results}

This chapter presents the empirical outcomes of our evaluation, focusing on search space reduction, memory behavior, and playing strength calibration. We begin with a summary table of the primary quantitative findings.

\begin{table}[htbp]
\centering
\caption[Key Findings: Quantitative Outcomes]{Key Findings: Quantitative Outcomes. Node Reduction measures tree size compression; Elo ladder gain represents relative playing strength changes fitted via Bradley-Terry; EBF is the Effective Branching Factor; Heap Used is V8 heap occupancy.}
\label{tab:key_findings}
\begin{tabularx}{\textwidth}{lX}
\toprule
\textbf{Finding} & \textbf{Value / Outcome} \\
\midrule
Peak Internal self-play rating & \textbf{1,400} (at search depth 6 in closed pool) \\
Depth-7 Regression & Drops to 1,365 Elo ($-$35) due to worker search timeout aborts \\
Largest Node Reduction & LMR: \textbf{66.3\%} node reduction (8,286 vs.\ 24,572 nodes at Depth 4) \\
Largest Elo Gain & Quiescence Search: $+$371 Elo ladder gain at Depth 3 \\
Positional Evaluation Gain & PSTs: $+$358.8 Elo ladder gain at Depth 2 \\
EBF Reduction & EBF drops from 9.22 (Depth 2) to 4.43 (Depth 8) \\
Memory Stability & 32.85\,MB -- 445.66\,MB heap usage across optimized runs \\
Transposition Hit Rate & 85.2\% peak hit rate with 500,000-entry FIFO cap \\
Statistical Significance & Only Quiescence Search accepted under SPRT \\
Stockfish Calibration & Scores 22.5\% vs.\ Stockfish Depth 3 ($-$215 relative depth-limited matchup differential, preliminary) \\
\bottomrule
\end{tabularx}
\end{table}



\clearpage

\section{Search Optimizations}

\subsection{Negamax Alpha-Beta Pruning}
We implement a Negamax formulation of alpha-beta pruning as our search foundation. Without pruning, game-tree traversal exhibits worst-case $O(b^d)$ complexity, expanding over 300,000 nodes at depth 4. Alpha-beta pruning restricts node expansions to 21,784 in our baseline, serving as the benchmark configuration against which subsequent optimizations are evaluated.

\subsection{Principal Variation Search (PVS)}
Principal Variation Search (PVS) refines search space compression by evaluating the first move with a full window, then probing subsequent alternatives with a zero-width window. Under our test conditions at depth 4, PVS yields a 5.2\% node reduction (from 8,739 to 8,286 nodes) and a 1.28× speedup. The performance gain remains modest due to overlap with LMR, although its efficiency increases at deeper search horizons where the principal variation is longer.

\section{Move Ordering Heuristics}
The efficiency of every pruning heuristic depends on move ordering. If optimal moves are evaluated early, the search window narrows rapidly, causing subsequent branches to fail low.

\subsection{MVV-LVA Capture Ordering}
Sorting captures by ``Most Valuable Victim -- Least Valuable Attacker'' (MVV-LVA) serves as the primary heuristic for ordering capture moves. By prioritizing tactical exchanges, MVV-LVA increases the probability of early beta cutoffs on capture branches. In our benchmarks, the complete search optimization stack (with all ordering and pruning enhancements active) compresses the search tree from 47,159 nodes in the baseline alpha-beta search (without ordering or search enhancements) to 8,286 nodes at depth 4 (Table~\ref{tab:node_counts}), representing an 82.4\% node reduction and an 8.33× speedup compared to this unordered baseline. This highlights how effective move sequencing enables the cumulative pruning heuristics in the engine to achieve their optimal theoretical efficiency.

\subsection{Killer Moves and History Heuristic}
To order quiet moves, Kronos uses a two-slot Killer Move cache per ply (storing quiet moves that triggered beta cutoffs in sibling nodes) and a History Table (tracking quiet moves that caused cutoffs globally, weighted by $\text{depth}^2$).

Enabling the History Heuristic yields a 6.6\% node reduction (from 8,869 to 8,286 nodes) at depth 4. While Killer Moves are measured to reduce search space, their individual playing-strength contribution remains statistically inconclusive under SPRT at the tested sample size, reporting a relative gain of $6 \pm 17$ internal Elo at depth 3. These ordering heuristics maintain list quality, allowing LMR to safely reduce search depths on low-priority branches.

\section{Search Extensions and Reductions}

\subsection{Late Move Reduction (LMR)}

\subsubsection{LMR Mechanism}
LMR is the most significant search reduction heuristic in Kronos. The intuition is straightforward: if moves are sorted well, the moves at the end of the list are very unlikely to be optimal. LMR searches these late moves at reduced depth ($d - 1$), treating them as unlikely candidates unless a reduced-depth probe returns a score above $\alpha$—in which case a full-depth re-search is triggered.

\subsubsection{LMR Experimental Result}
The ablation data quantifies this clearly. Enabling LMR yields a \textbf{66.3\%} node reduction (from 24,572 down to 8,286 nodes), resulting in a 4.17× speedup (execution speed drops to 0.24× when disabled). This is by far the largest single-heuristic contribution in our search ablation.

\subsubsection{Interpretation and Ordering Dependency}
The mechanism is simple: without LMR, every move at every node must be searched to full depth. With LMR, the majority of the search tree is explored at shallower depth. The safety of this reduction depends entirely on move ordering quality—poor ordering causes good moves to appear late in the list and be searched at reduced depth, missing the optimal move. This creates a tight dependency between LMR and the ordering heuristics described above.


\input{tables/node_counts}

\subsection{Null Move Pruning (NMP)}
NMP prunes search branches where passing the turn to the opponent still results in a score exceeding $\beta$, indicating that the position is sufficiently dominant to guarantee a cutoff. We configure NMP for search depths $d \ge 5$ using a constant reduction factor $R = 2$.

At depth 4, under the final optimized configuration ($d \ge 5$), NMP is inactive, resulting in a 0.0\% node reduction (8,286 nodes) and a 1.00x relative speedup. At deeper search plies ($d \ge 5$), NMP becomes active, pruning dominant branches early and avoiding expensive lower-ply move generation and evaluation calls on pruned sub-branches. The significance of NMP scales with search depth, proving critical for staying within time controls at depth 6.

\subsection{Quiescence Search}
Quiescence search mitigates the horizon effect by extending the search beyond the nominal depth limit for unstable, tactical positions containing legal captures. This prevents the engine from returning misleading static evaluations mid-sequence.

Disabling quiescence search at depth 3 decreases playing strength from 1,387 to 1,016 in internal Bradley-Terry self-play rating. This drop of 371 internal Elo represents the largest single-heuristic effect in our ablation study, highlighting the necessity of quiescence search in mitigating tactical blindness.

\subsection{Evaluation Function: Piece-Square Tables}
Substituting a material-only evaluation with Piece-Square Tables (PSTs) provides an internal rating improvement of 358.8 Elo at depth 2. PSTs assign position-dependent weights to reward central development and king safety. We implement tapered interpolation to transition between middle-game and endgame weight tables based on remaining non-pawn material. Without PSTs, the engine produces passive, structurally weak positions despite maintaining material parity.

\subsection{Memory Engineering and GC Stability}
Table~\ref{tab:memory_profile} documents memory usage across configurations. Specifically, the early unoptimized configurations (EXP-5-TRANSPOSITION and EXP-6-ITERATIVE) utilized a Map-based transposition table capped at a default soft limit of 200,000 entries. Due to the high object allocation overhead of V8 Map entries (each wrapping a complex key-value structure in heap memory), even this entry limit resulted in heap exhaustion, with RSS exceeding 2\,GB and 3.4\,GB respectively. These early outcomes motivated the implementation of a strict size-bounded \texttt{Map}-based transposition table with FIFO eviction (capped at 500,000 entries), eliminating unbounded memory growth while preserving O(1) amortized lookup and insertion performance.

\input{tables/memory_profile}

With the FIFO capacity limit active, the V8 heap occupancy remains bounded between 32.85\,MB and 445.66\,MB across final optimized tournament runs (depending on search depth and match count), with single-game telemetry profiling showing active heap usage as low as 24.98\,MB. This simple policy restricts the active cache memory to approximately 30\,MB while maintaining an 85.2\% hit rate. Although the transposition table at 72,288 entries contributes less than 15\,MB of direct data overhead, the remaining heap footprint (reaching a peak of 445.66\,MB) is occupied by transient V8 system objects, including evaluation history stacks, move ordering priority queues, the JIT compiler's internal type-feedback buffers, and accumulated string and move objects generated internally by \texttt{chess.js} that await garbage collection sweeps. This hit rate is sustained because chess searches frequently evaluate transposed board states during middle-game traversal.

\section{Search Depth Scaling and NPS Throughput}
Table~\ref{tab:cross_depth_elo} documents per-move search throughput and playing strength scaling across depths 1 through 7.

\input{tables/cross_depth_elo}

The internal Bradley-Terry self-play rating scales from 1,015 (estimated) at depth 1 to a peak internal self-play rating of 1,400 Elo at depth 6. The ratings for depths 1 and 2 are estimated by anchoring the Bradley-Terry self-play model to the baseline minimax performance (anchored at 1000) and extrapolating the rating curve using the win-loss ratios of a small paired matchup series against depth 3. At depth 5, the rating plateaus at 1,383 Elo, suggesting that the additional search ply does not consistently find better moves within the configured tournament time controls.

The performance regression at depth 7 ($-$35 Elo) is caused by single-threaded latency constraints. Telemetry logs indicate that the worker thread exceeded the move time limit (forcing a fallback to the depth-6 results) on average for 42.5\% of the moves during depth-7 matches. When search time at depth 7 exceeds the tournament time limit, the engine returns the best move from the last fully completed iterative deepening iteration (depth 6) rather than the incomplete depth-7 results. This timeout behavior introduces a performance ceiling, indicating that single-threaded JavaScript search workers cannot reliably sustain depth-7 calculations under standard time controls.

The Effective Branching Factor (EBF) decreases from 9.22 at depth 2 to 4.43 at depth 8 (Figure~\ref{fig:effective_branching_factor}). This contraction shows that our pruning heuristics become proportionally more effective as the search depth increases, since deeper horizons provide more contextual information for move ordering and reduction decisions.

\section{Stockfish Reference Calibration}
To provide a relative external reference for Kronos's internal ratings, we benchmarked the final Depth 6 engine configuration against three depth-limited Stockfish 18 configurations. These calibration matches measure the relative depth-limited Stockfish matchup differential---representing performance against fixed-depth references rather than an absolute Elo anchor. Direct mapping to FIDE or CCRL rating scales is not claimed.

\input{tables/stockfish_calibration}


Over the calibration matches, Kronos scores 32.5\% against Stockfish Depth 1 (400 games), 35.0\% against Stockfish Depth 2 (280 games), and 22.5\% against Stockfish Depth 3 (400 games). These outcomes correspond to relative depth-limited matchup differentials of $-127 \pm 25$ Elo, $-108 \pm 25$ Elo, and $-215 \pm 29$ Elo, respectively. This reports a systematic strength gap against depth-limited Stockfish, scaling with opponent search horizon.

\textbf{Important caveat:} Stockfish 18 relies heavily on NNUE (Efficiently Updatable Neural Network) evaluation, which propagates learned features across deep search plies. Restricting Stockfish to depths 1--3 non-linearly degrades its effective playing strength, because the NNUE evaluation benefits from deep search propagation and advanced pruning heuristics that only activate at deeper plies. Consequently, depth-limited Stockfish does not scale linearly with search depth, and these matchup differentials should not be interpreted as anchors to standard Elo rating pools.

\subsection{Statistical Significance of Heuristic Gains}
Sequential Probability Ratio Tests (SPRT) evaluate the statistical significance of these playing strength gains under Wald boundaries $[-2.944, 2.944]$.

\input{tables/statistical_validation}

Only quiescence search (LLR = 75.433) crosses the upper SPRT boundary to be accepted ($H_1$), indicating a tactical strength improvement. The other pairwise tournament runs (such as MVV-LVA move ordering vs.\ alpha-beta only, LLR = 0.859, and iterative deepening vs.\ transposition table caching, LLR = $-$0.171) remain inconclusive under SPRT, indicating that their playing strength differences are small and do not cross the SPRT threshold under the configured sample sizes.

To isolate and verify the baseline rating profile of each engine configuration without the confounding factor of execution timeouts, we downshift search horizons to depth 3 for the self-play round-robin tournament in Table~\ref{tab:tournament_results}. This tournament results summary at depth 3 highlights the influence of quiescence search: the full engine variant scores 90.0\% (1,387 internal self-play rating) across 60 games in the self-play tournament, whereas the variant without quiescence search scores 45.0\% (1,016 internal self-play rating). Note that this rating of 1,387 Elo represents strength fitted within a depth-3 configuration ablation pool (anchored at Baseline Minimax Depth 3 = 1000). It is mathematically distinct from the 1,295 Elo rating reported for Depth 3 in Table~\ref{tab:cross_depth_elo}, which represents strength fitted within a cross-depth pool of final engine variants (anchored at Baseline Minimax Depth 1 = 1000). Due to different player mixes and anchors, the relative rating scales are pool-dependent.

\input{tables/tournament_results}

We observe that \textit{Baseline Minimax}, \textit{Alpha-Beta Only}, \textit{Move Ordering (MVV-LVA)}, \textit{Killer Moves}, and \textit{Transposition Table} produce nearly identical win/draw/loss records in Table~\ref{tab:tournament_results}. This is a direct consequence of search equivalence under fixed-depth constraints: all five configurations are mathematically equivalent formulations of the minimax algorithm. Since Alpha-Beta pruning, move ordering heuristics, and transposition table caching preserve the root node evaluation value, they do not alter the engine's move selection under normal conditions. 

The slight rating improvement (+5 Elo) and variance in draw/loss records for the optimized configurations are explained by systems-level execution limits. In slower, unoptimized configurations (\textit{Baseline Minimax} and \textit{Alpha-Beta Only}), search latency spikes occasionally exceed the Web Worker execution time limits, forcing the search to abort early and return a default move. The optimized configurations avoid these timeouts, converting occasional losses into draws. In contrast, the \textit{Full Kronos Engine} achieves a dramatic rating increase to 1,387 Elo by integrating quiescence search, which alters move selections by resolving the horizon effect.

Figures~\ref{fig:search_pipeline}--\ref{fig:memory_usage} visualize the search pipeline structure, node growth, strength scaling, Stockfish calibration curve, EBF reduction, and memory profile.

\input{figures/search_pipeline}
\input{figures/node_growth}
\input{figures/strength_vs_depth}
\input{figures/stockfish_calibration}
\input{figures/effective_branching_factor}
\input{figures/memory_usage}


\section{Game-Level Computational Overhead}
To evaluate the aggregate computational overhead incurred across full matches (rather than single-move searches), Table~\ref{tab:search_cost} details the average nodes visited, total search time, and resulting aggregate NPS achieved per game across depths 2 through 5. As depth scales, the average nodes searched per game increases exponentially from 5,501 nodes at depth 2 to over 1.2\,million nodes at depth 5. This exponential growth is mirrored in the time scaling, showing that depth-5 tournament games require substantial computational execution time (131.87\,seconds per game) on V8. The aggregate NPS values (varying between 2,515 and 10,423 NPS) reflect the dynamic JIT optimizations and transposition cache hit rates across complete game lifecycles.

Note that the game-level NPS values in Table~\ref{tab:search_cost} differ from the per-move NPS values reported in Table~\ref{tab:cross_depth_elo}. While Table~\ref{tab:cross_depth_elo} measures search throughput for isolated, single-move evaluations, Table~\ref{tab:search_cost} reports the aggregate throughput computed over complete game lifecycles. This aggregate throughput is heavily influenced by dynamic runtime effects, including V8 JIT warmup optimization phases, accumulated transposition table cache hits in mid-game positions, and varying tactical branches across consecutive moves. Specifically, the non-monotonic scaling—where Depth 3 spikes to 10,423 NPS before dropping to 2,937 NPS at Depth 4—is an empirical consequence of cache-to-tree ratios and runtime memory management. At Depth 3, the transposition table cache hits accumulate rapidly and fit entirely within V8's cache limits with negligible memory sweeps. At Depth 4, the exponential tree expansion triggers transposition table cache evictions and increases allocation pressure, inducing minor garbage collection cycles that degrade the aggregate throughput.

\input{tables/search_cost}
