\chapter{Future Work}
\label{ch:future_work}

Several opportunities exist to extend the performance and capabilities of Kronos:
\begin{itemize}
    \item \textbf{Custom Move Generator and Board Representation:} The current engine delegates move generation to the third-party \texttt{chess.js} library, which is the dominant throughput bottleneck. Replacing \texttt{chess.js} with a custom typed-array bitboard generator would eliminate the mailbox overhead and is expected to increase NPS by an order of magnitude, even within JavaScript.
    \item \textbf{Dual-32-bit Zobrist Hashing:} The current Zobrist hashing implementation uses JavaScript \texttt{BigInt} primitives for 64-bit keys. A standard optimization in JavaScript chess engines is to represent each 64-bit hash as two 32-bit integers, enabling native bitwise XOR operations without \texttt{BigInt} object overhead. This migration would reduce per-node hashing cost on the hot search path.
    \item \textbf{WebAssembly Migration:} Migrating the core search loop and board state mutations to WebAssembly (Wasm) represents a natural next step. While V8 JIT compilation offers optimized execution speed, WebAssembly would eliminate JIT warm-up times and compile-timing variations, providing consistent execution speeds.
    \item \textbf{Neural Network Evaluations (NNUE):} Replacing the classical static evaluation function with an Efficiently Updatable Neural Network (NNUE) would significantly improve evaluation quality. Future work will explore implementing NNUE inference inside web runtimes, utilizing WebGL/WebGPU or Wasm vector instructions.
    \item \textbf{Parallel Search Implementations:} The current Kronos worker architecture is single-threaded. We plan to explore parallel search algorithms, such as Lazy SMP or Principal Variation Split, utilizing shared memory abstractions (\texttt{SharedArrayBuffer}) and multi-worker execution pools to scale search performance on multi-core systems.
    \item \textbf{Caching, Tablebases, and Variance Heuristics:} Integrating endgame tablebase caching (such as Syzygy tablebases) and expanding our opening books would improve playing strength in the opening and endgame phases. Additionally, to mitigate the high draw rates (66.7\%--70.0\%) observed in self-play matches between similar variants, we plan to explore introducing temperature-scaled randomness or multi-PV (Principal Variation) selections in the opening plies to force more variance and reduce these deterministic draw loops.
\end{itemize}
