\chapter{Limitations and Threats to Validity}
\label{ch:limitations}

This chapter details the throughput limitations of the Kronos search core and outlines potential threats to the validity of our experimental findings, alongside mitigation steps.

\section{System Limitations}
The primary limitation of Kronos is its search throughput, which is constrained by two architectural factors. First, the engine delegates board representation and move generation to the third-party \texttt{chess.js} library, which relies on a mailbox (0x88) board representation with complex string parsing and object validation on internal code paths. Unlike native C/C++ engines that utilize 64-bit integer bitboards updated via bitwise CPU operations, \texttt{chess.js} operates through object-based wrappers. Replacing \texttt{chess.js} with a custom typed-array bitboard generator would be the most impactful single optimization for throughput. Second, the Zobrist hashing subsystem uses JavaScript \texttt{BigInt} primitives for 64-bit key operations. While \texttt{BigInt} correctly preserves 64-bit precision, it incurs measurable overhead compared to the dual-32-bit integer representation commonly used in production JavaScript chess engines (e.g., representing a 64-bit hash as two 32-bit integers with native bitwise XOR operations). Together, these factors limit throughput to approximately 5,000 NPS (Table~\ref{tab:cross_depth_elo}).

Consequently, the engine cannot search beyond 6 plies under standard time controls. Furthermore, the search core is single-threaded, preventing it from utilizing multi-core processor architectures.

\section{Threats to Validity}

\subsection{Internal Validity}
Internal validity threats concern execution variables that can affect timing and search statistics. In managed runtime environments, background operating system activity, JIT compiler warming, and garbage collection cycles introduce timing variations that can skew search latency benchmarks.

To mitigate these timing variations, all tournaments were executed on isolated hardware (Intel Core i5-1235U CPU, 16\,GB RAM) with background processes disabled. JIT warming was managed by executing initial dummy games to ensure V8 compiler optimization completed before recording metrics. However, our benchmarks are inherently tied to this hardware profile and the Node.js v20.11.0 runtime (Google V8 Engine). Variations in clock speeds, memory latency, and JIT optimization heuristics across other platforms (such as Bun, Deno, or older V8 releases) represent unmeasured threats to internal validity.

\subsection{External Validity}
External threats limit the generalizability of our findings. First, the engine's relative rating scaling and pruning efficiencies were evaluated using a specific opening library (\texttt{positions.json}) limited to 50 FEN positions. Results could differ under rare, highly unbalanced, or theoretically complex opening lines. We addressed color and opening bias by enforcing paired color-swapped matches, but the restricted size of the openings book remains a limitation.

Second, Kronos lacks several components common in modern tournament chess engines, such as endgame tablebases (e.g., Syzygy tablebases) and Neural Network evaluations (NNUE). The absence of tablebases increases draw rates in simplified endgames, as the engine cannot resolve forced mates or draws instantly. The absence of NNUE limits the positional and tactical accuracy of evaluations, restricting the engine's play to classical heuristic weightings (Piece-Square Tables).

\subsection{Construct and Conclusion Validity}
Construct validity threats relate to our playing strength metrics. The internal playing strength ratings reported in this work represent relative measurements within a closed self-play pool of Kronos engine variants, not absolute FIDE-style ratings. To provide an external reference, we benchmarked Kronos against depth-limited Stockfish 18 configurations. Although these Stockfish calibration matches provide a preliminary external reference anchored against fixed-depth configurations, depth-limited Stockfish does not degrade linearly: its NNUE evaluation architecture is designed for deep search propagation, and restricting Stockfish to depths 1--3 non-linearly cripples its evaluation accuracy. These matchup differentials therefore cannot be directly mapped to standard FIDE or CCRL Elo scales, and projected ratings for deeper Stockfish configurations are speculative estimates that should not be treated as measured empirical data.

Additionally, statistical conclusion validity is bounded by the deterministic nature of the search core. Because Kronos uses strictly deterministic root move selection (no temperature scaling or stochastic sampling), the effective degrees of freedom in any matchup are bounded by the opening book size: $2 \times |\text{opening book}| = 100$ independent game outcomes per variant pair. Sample sizes beyond this bound do not contribute new statistical information.

Finally, the 100 to 400 game tournaments executed at depth 4 provide standard error margins of approximately $\pm 17$ to $\pm 34$ Elo. However, the draw rates observed during self-play matches between similar variants (exceeding 66.7\% to 70.0\%) significantly reduce the statistical information gain per game, widening the confidence intervals. While this is sufficient to confirm significance for major structural changes (like quiescence search or move ordering), smaller rating differentials require much larger game samples to resolve conclusively. We address this limitation by reporting the full LLR status and trinomial confidence intervals for all tournament outcomes, explicitly identifying cases where results remain inconclusive.

\subsection{Dependency Fragility and Reproducibility}
A critical engineering threat to long-term reproducibility is Kronos's reliance on the internal, private API of the third-party \texttt{chess.js} library. To eliminate garbage collection overhead in the move generation loop, the custom wrapper class \texttt{ChessInstanceClone} bypasses the library's public interfaces and directly invokes private methods (e.g., \texttt{\_moves()}, \texttt{\_makeMove()}, and \texttt{\_undoMove()}). While this achieves GC neutrality in the current workspace (Node.js environment with \texttt{chess.js} v1.4.0), it introduces severe dependency fragility. Future minor or major updates to \texttt{chess.js} that alter or remove these internal private functions will break the search core. To mitigate this risk, future work must transition from third-party wrappers to an independent, custom-built move generator.

